---
title: "Use logit regression models to predict Titanic passenger survival"
author: "Fiona, Fang Jiaying"
date: '2022-11-9'
output: github_document
---

# Preparation  

## Packages  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rstudioapi::getActiveDocumentContext
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
```

```{r}
pacman::p_load(tidyverse, haven, stargazer, tidymodels, ranger, modelsummary, dplyr, modeldata, pscl, pROC)
```


## Data  

We load the Titanic data.  

```{r}
td <- read.csv("titanic.csv")
```

We need to clean up the data to prepare for data analysis.  

Let's first explore the target variables (survived).  

```{r}
with(td, table(survived, useNA = "ifany"))
```

There is no missing value in the dependent variable.  

Then, {tidymodels} require that the dependent variable is stored as a factor variable (categorical variable) when we do classification. Therefore, we should change the data type.  

```{r}
td <- td |> 
  mutate(survived_fac = factor(survived))

with(td, table(survived, survived_fac, useNA = "ifany"))
```

Let’s check if there’s any missing value in data test.
```{r}
colSums(is.na(td))
```


Drop observations with missing values.  

```{r}
td.nna <- td |> 
        dplyr::select(survived_fac, pclass, age, child, old, sex, sibsp, parch, alone, fare, cherbourg, queenstown, southampton) |> 
        na.omit()

colSums(is.na(td.nna))
```

Then, we want to use logit regression models to predict Titanic passenger survival.


# Logit Models

Run the logit models to view the results

```{r}
td.nna <- td.nna |> 
  mutate(pclass = factor(pclass))

mod1 <- glm(survived_fac ~., family = binomial(link = 'logit'), data = td.nna)

stargazer(mod1, type = "text",
          dep.var.labels=c("Likelihood of survival"),
          covariate.labels=c("pclass2",
                             "pclass3",
                             "Age",
                             "Child", 
                             "Old",
                             "Woman",
                             "Number of siblings or spouses",
                             "Number of parents or children",
                             "Alone",
                             "Fare",
                             "Departure from Cherbourg",
                             "Departure from Queenstown",
                             "Departure from Southampton"))
```


From the result, we can see that pclass, age, child, woman, Alone, cherbourg, sibspp, and queenstown are  significant features for predicting survival outcome. For instance, one-unit increase in age will decrease the log odd of survival by 0.023.     

# Models: logit regression

Then, we try to figure out the best model. 

```{r}
td.nna <- td.nna |> 
  mutate(pclass = factor(pclass))

fit1 <- logistic_reg() |> 
  fit(survived_fac ~ pclass + age + child + old + 
        sex + sibsp + parch + alone + fare + 
        cherbourg + queenstown + southampton, 
      data = td.nna)

fit2 <- logistic_reg() |> 
  fit(survived_fac ~ pclass + child + sex + sibsp, 
      data = td.nna) 

fit3 <- logistic_reg() |> 
  fit(survived_fac ~ pclass + child + sex + sibsp +  
                  age + cherbourg + alone, 
      data = td.nna) 

fit4 <- logistic_reg() |> 
  fit(survived_fac ~ pclass + child + sex + sibsp +  
                  age + cherbourg + alone+ 
               queenstown, 
      data = td.nna) 

fit5 <- logistic_reg() |> 
  fit(survived_fac ~ pclass + child + sex + sibsp +  
                  age  + cherbourg + alone + 
                  queenstown + parch, 
      data = td.nna) 


modelsummary(list(fit1, fit2, fit3, fit4, fit5),
             coef_rename = c(
               "(Intercept)" = "Intercept", 
               "pclass" = "",
               "age" = "Age",
               "childChild" = "Child", 
               "old" = "Old",
               "sexWoman" = "Woman",
               "sibsp" = "Number of siblings or spouses",
               "parch" = "Number of parents or children",
               "alone" = "Alone",
               "fare" = "Fare",
               "cherbourg" = "Departure from Cherbourg",
               "queenstown" = "Departure from Queenstown",
               "southampton" = "Departure from Southampton"))

```

# Cross validation

We use 5-fold cross validation (outof-sample prediction) to decide the best model.  

```{r}
set.seed(12345)
cv_fold <- vfold_cv(td.nna, v = 5)
```

```{r}
fit.1_cv <- logistic_reg() |> 
  fit_resamples(survived_fac ~ pclass + age + child + old + 
        sex + sibsp + parch + alone + fare + 
        cherbourg + queenstown + southampton, 
                resamples = cv_fold)

collect_metrics(fit.1_cv)
```

```{r}
fit.2_cv <- logistic_reg() |> 
  fit_resamples(survived_fac ~ pclass + child + sex + sibsp, 
                resamples = cv_fold)

collect_metrics(fit.2_cv)
```


```{r}
fit.3_cv <- logistic_reg() |> 
  fit_resamples(survived_fac ~ pclass + child + sex + sibsp +  
                  age + cherbourg + alone, 
                resamples = cv_fold)

collect_metrics(fit.3_cv)
```


```{r}
fit.4_cv <- logistic_reg() |> 
  fit_resamples(survived_fac ~ pclass + child + sex + sibsp +  
                  age + cherbourg + alone +
                  queenstown, 
                resamples = cv_fold)

collect_metrics(fit.4_cv)
```

```{r}
fit.5_cv <- logistic_reg() |> 
  fit_resamples(survived_fac ~ pclass + child + sex + sibsp +  
                  age  + cherbourg + alone + 
                  queenstown + parch, 
                resamples = cv_fold)

collect_metrics(fit.5_cv)
```

The mean of roc_auc of fit 5 is the largest here. Thus, it is the best model to choose.  


# Evaluating performances

Calculate predicted probabilities.  

```{r}
pred_logit_1 <- predict(fit1, td.nna, type = "prob")
pred_df_logit_1 <- data.frame(survived_fac = td.nna $ survived_fac, pred_logit_1) 

pred_logit_2 <- predict(fit2, td.nna, type = "prob")
pred_df_logit_2 <- data.frame(survived_fac = td.nna $ survived_fac, pred_logit_2) 

pred_logit_3 <- predict(fit3, td.nna, type = "prob")
pred_df_logit_3 <- data.frame(survived_fac = td.nna $ survived_fac, pred_logit_3) 

pred_logit_4 <- predict(fit4, td.nna, type = "prob")
pred_df_logit_4 <- data.frame(survived_fac = td.nna $ survived_fac, pred_logit_4) 

pred_logit_5 <- predict(fit5, td.nna, type = "prob")
pred_df_logit_5 <- data.frame(survived_fac = td.nna $ survived_fac, pred_logit_5) 
```


Create a ROC figure that compares the five models

```{r}
roc_logit_1 <- roc(pred_df_logit_1 $ survived_fac ~ 
                   pred_df_logit_1 $.pred_1 + 
                   pred_df_logit_2 $.pred_1, ci = TRUE)

roc_logit_2 <- roc(pred_df_logit_1 $ survived_fac ~ 
                   pred_df_logit_1 $.pred_1 + 
                   pred_df_logit_3 $.pred_1, ci = TRUE)

roc_logit_3 <- roc(pred_df_logit_1 $ survived_fac ~ 
                   pred_df_logit_1 $.pred_1 + 
                   pred_df_logit_4 $.pred_1, ci = TRUE)

roc_logit_4 <- roc(pred_df_logit_1 $ survived_fac ~ 
                   pred_df_logit_1 $.pred_1 + 
                   pred_df_logit_5 $.pred_1, ci = TRUE)
```

ROC curves together:  

```{r, tidy=F}
ggroc(roc_logit_1, 
      legacy.axes = TRUE, aes = c("linetype")) + 
        geom_abline(color = "dark grey", size = 0.5) + 
        labs(x = "False positive rate", 
             y = "True positive rate",
             linetype="Models") + 
        scale_linetype_discrete(labels=c("Model1", "Model2")) + 
  theme_bw() + 
        theme(legend.position=c(0.8,0.2))

ggroc(roc_logit_2, 
      legacy.axes = TRUE, aes = c("linetype")) + 
        geom_abline(color = "dark grey", size = 0.5) + 
        labs(x = "False positive rate", 
             y = "True positive rate",
             linetype="Models") + 
        scale_linetype_discrete(labels=c("Model1", "Model3")) + 
  theme_bw() + 
        theme(legend.position=c(0.8,0.2))

ggroc(roc_logit_3, 
      legacy.axes = TRUE, aes = c("linetype")) + 
        geom_abline(color = "dark grey", size = 0.5) + 
        labs(x = "False positive rate", 
             y = "True positive rate",
             linetype="Models") + 
        scale_linetype_discrete(labels=c("Model1","Model4")) + 
  theme_bw() + 
        theme(legend.position=c(0.8,0.2))


ggroc(roc_logit_4, 
      legacy.axes = TRUE, aes = c("linetype")) + 
        geom_abline(color = "dark grey", size = 0.5) + 
        labs(x = "False positive rate", 
             y = "True positive rate",
             linetype="Models") + 
        scale_linetype_discrete(labels=c("Model1","Model5")) + 
  theme_bw() + 
        theme(legend.position=c(0.8,0.2))
```

Test if the AUCs are different:  

```{r}
roc.test(roc1 = roc_logit_1[[1]], 
         roc2 = roc_logit_1[[2]], method = "delong")

roc.test(roc1 = roc_logit_2[[1]], 
         roc2 = roc_logit_2[[2]], method = "delong")

roc.test(roc1 = roc_logit_3[[1]], 
         roc2 = roc_logit_3[[2]], method = "delong")

roc.test(roc1 = roc_logit_4[[1]], 
         roc2 = roc_logit_4[[2]], method = "delong")
```

We can see from above that AUCROC of Model 5 is the largest. Thus, it is the best model to choose. 




